Vision Paper: Efficient Spatio-Temporal Geo-Statistics for Exploratory Data Analysis

### 2018-05-24 Workshop_GIScience_BigData_Vision.tex
updated paper

\begin{abstract}
Hot spot analysis is essential for geo-statistics.
It supports decision making by detecting points
as well as areas of interest in comparison to their
neighbourhood. However, these methods are dependent
on different parameters, ranging from the resolution
of the study area to the size of their neighbourhood.
This dependence can lead to instabilities of the detected
hotspots, where the results can highly vary between
different parameters. A decision maker can therefore ask
how valid the analysis actually is.
%
In this study, we examine the impact of key parameters
on the stability of the hotspots, namely the size of
the neighbourhood, the resolution and the size of the
study area, as well as the influence of the ratio between
those parameters.
%
We compute the hotspots with the well known Getis-Ord ($G^*$)
statistic as well as its modification, the \emph{Focal $G^*$} statistic.
We measure the stability of the hotspot analysis using
a recently introduced \emph{stability of hotspots} metric (SoH)
and compare the results to intuitive visual analysis.
%
We evaluate the results on real world data with the well-known
yellow cab taxi data set from New York, Manhattan.
Our results indicate a negative impact on the stability with
a reduction of the size of the neighbourhood as well as
a reduction of the size of the study area, regardless of
the resolution.
\end{abstract}

@@

But, similar to a cluster analysis, the identified hotspots highly depend on 
the parametrization of the detection method. The identified areas as well as 
their shape can vary highly. This volatility can lead to a decrease in trust in 
the result or in suboptimal allocations of scarce resources. Therefore, it is 
necessary to measure and evaluate the stability of a hotspot analysis as well 
as the different parametrizations.

In our initial work~\cite{SoH-GI-Forum}, we introduced a method to measure the 
stability of hotspots, the \emph{stability of hotspots} metric (SoH) and showed 
its use on the basis of temperature data.
Here, we build upon that work and examine the impact of the 
different instantiations of the most typical parameters in more detail. We use 
the well known 
Getis-Ord statistic \cite{Ord.1995}, the standard $G^*$, and a modification of 
this statistic, the Focal~$G^*$~\cite{SoH-GI-Forum}.
Those parameters examined are
\begin{inparaenum}[(1)]
   \item size of the study area given as a focal matrix,
   \item pixel resolution i.e. aggregation level and 
   \item the size of the neighbourhood given as a weight matrix.
\end{inparaenum}
By varying over these parameter, we can compare the stability for all possible 
combinations and isolate the effect of a single parameter.
We evaluate the metrics on the well-known dataset of New York Taxi drives to 
show the applicability to real world use cases as well as to foster 
reproducibility.

\section{State of the Art}

@@

\subsection{Quality of Clustering}
The problem of assessing the quality in unsupervised learning is well known. In the case of the k-mean algorithm, the quality of the clustering is mostly dependent on the value of the \emph{k} and a miss-specification can lead to highly irregular clusters. In a simple 2D clustering, they can be easily recognized by visual analysis, but in higher dimensionality, this is impossible. One method, to measure the quality of such a clustering is the compactness of the clusters, see e.g. \cite{CompactnessDataClustering}. This enables the comparison between different clusters. Another possibility is the Silhouette Coefficient by Kaufman et Rousseeuw 1990. This metric measures the similarity of objects in a cluster in comparison to other clusters. For density based clustering, e.g. for DBSCAN~\cite{Ester96adensity-based}, OPTICS~\cite{Ankerst:1999:OOP:304181.304187} gives a simple method to tune the essential parameter for this clustering. This is only a small overview of methods to influence and measure the quality of different clustering methods. But it shows that this problem is not easily solved and dependent on the chosen algorithm. To our knowledge, there does not exist a method to overall measure the stability of a clustering.

\subsection{Hot Spot Analysis}
The goal of hotspot analysis is the detection of interesting areas as well as patterns in spatial information. One of the most fundamental approach is Moran´s I~\cite{MoranI}. There it is tested whether or
not a spatial dependency exists. This gives the information on global
dependencies in a data set. Upon this hypothesis test several geo-statistical
tests are based. The most well known are the Getis-Ord statistic~\cite{Ord.1995}
and LISA~\cite{Anselin.1995}. In both cases the general, the global statistic of
Moran´s I is applied in a local context. The goal is to detect not only global
values, but instead to focus on local hotspots and to measure the significance
of those local areas. A more in depth overview of methods to identify and visualize spatial patterns and areas of interest can be found in \cite{shekhar2011identifying}.

\section{Stability of Hot Spot Analysis}
\label{sec:Metric}

Existing methods for determining hot spots are dependent on the parametrization
of the weight matrix as well as on the size of the study area. Intuitively,
increasing the size of a weight matrix has a "blurring" effect on the raster
(Figure~\ref{fig:BlurExample:a}) whereas decreasing the size can be 
seen as a form of "sharpening" (Figure~\ref{fig:BlurExample:b}).

%TODO: add image showing and blurring/sharpening effects

%As we can see, hot spots are oftentimes disappearing or appearing unrelated to
%previously found hot spots. While these computations indeed show hot spots and
%the results are correct, they lack stability.

For a data analyst, when exploring the data interactively by choosing different
filter sizes (weight matrices) or point aggregation strategies (pixel sizes), it
is important that the position and size of a hot spot changes in a predictable
manner. We formalize the intuition in our stability metric.

%
%Consider the real-world example depicted in Figure~\ref{fig:TempMaps}.
%%TODO: thermal flight -> NY taxi
%The temperature map of a morning thermal flight dataset 
%(Figure~\ref{fig:TempMaps:a}) has been processed using $G^*$.
%
%%TODO: thermal flight -> NY taxi
%%TODO: need two images - G* and FocalG*
\begin{figure}[htp]
  \centering
  \begin{subfigure}{.3\linewidth}
    \caption{Small matrix}
    \includegraphics[width=\linewidth]{images/gen-raw-blur-gstar-2}
    \label{fig:BlurExample:a}
  \end{subfigure}
  \begin{subfigure}{.3\linewidth}
    \caption{Large matrix}
    \includegraphics[width=\linewidth]{images/gen-raw-blur-gstar-8}
    \label{fig:BlurExample:b}
  \end{subfigure}
  \caption{
    Example of G* statictics computed for the same raster with different weight 
    matrix sizes.
  }
\end{figure}

We define a hot spot found in comparably more coarse resolutions as
\emph{parent} (larger weight matrix or larger pixel size) and in finer
resolutions as \emph{child} (smaller weight matrix or smaller pixel size).

Stability assumes that every parent has at least one child and that each child
has one parent. For a perfectly stable interaction, it can be easily seen that
the connection between parent and child is a injective function and between
child and parent a surjective function. To measure the stability,
we propose a metric called the \emph{Stability of Hot spot} (SoH). It measures
the deviation from a perfectly stable transformation between the parent and 
chaild rasters (like frames of an animation).

In its downward property (from parent to child, injective) it is defined as:
\begin{equation}
  \label{eq:SoH-down}
  \SOHDOWN
    = \frac{ParentsWithChildNodes}{Parents}
    = \frac{|Parents \cap Children|}{|Parents|}
\end{equation}
And for its upward property (from child to parent, surjective):
\begin{equation}
  \label{eq:SoH-up}
  \SOHUP
    = \frac{ChildrenWithParent}{Children}
    = 1 - \frac{|Children - Parents|}{|Children|}
\end{equation}
where $ParentsWithChildNodes$ is the number of parents that have at least one 
\emph{child}, $Parents$ is the total number of \emph{parent}, 
$ChildrenWithParent$ is the number of children and $Children$ as the total 
number of children.
The SoH is defined for a range between 0 and 1, where 1 represents a perfectly 
stable transformation while 0 would be a transformation with no stability at 
all.
If $|Children|=0$ or $|Parents|=0$ SoH is 0.

%\section{Focal Getis-Ord}
%\label{sec:FocalGetisOrd}
%
%In paper~\cite{SoH-GI-Forum}, we have defined the measure Focal G*.
%We use the notation \FOCAL{R}{M}{op} to denote a focal
%operation $op$ applied on a raster $R$ with a focal window determined by a
%matrix $M$. This is rougly eqivalent to a command \verb|focal(x=R, w=M, 
%fun=op)|
%from package \emph{raster} in the R programming language \cite{cran:raster}.

%\begin{definition}[$G^*$ function on rasters]
%  \label{def:GenericGetisOrdFunc}
%  
%  The function $G^*$ can be expressed as a raster operation:
%  
%  \begin{displaymath}
%    G^*(R, W, st) =
%    \frac{
%      \FOCAL{R}{W}{sum} - M*\sum_{w \in W}{w}
%    }{
%      S \sqrt{
%        \frac{
%          N*\sum_{w \in W}{w^2} - (\sum_{w \in W}{w})^2
%        }{
%          N - 1
%        }
%      }
%    }
%  \end{displaymath}
%  \noindent where:
%  \begin{itemize}
%    
%    \item $R$ is the input raster.
%    
%    \item $W$ is a weight matrix of values between 0 and 1.
%    
%    \item $st = (N, M, S)$ is a parametrization specific to a particular 
%version
%    of the $G^*$ function. (Def.\vref{def:StdGetisOrdVersion} and
%    \vref{def:FocalGetisOrdVersion}).
%    
%  \end{itemize}
%\end{definition}
%
%\begin{definition}[Standard $G^*$ parametrization]
%  \label{def:StdGetisOrdVersion}
%  
%  Computes the parametrization $st$ as global statistics for all pixels in the
%  raster $R$:
%  
%  \begin{itemize}
%    \item $N$ represents the number of all pixels in $R$.
%    \item $M$ represents the global mean of $R$.
%    \item $S$ represents the global standard deviation of all pixels in $R$.
%  \end{itemize}
%\end{definition}
%
%\begin{definition}[Focal $G^*$ parametrization]
%  \label{def:FocalGetisOrdVersion}
%  
%  Let $F$ be a boolean matrix such that: $all(dim(F) \geq dim(W))$.
%  This version uses focal operations to compute per-pixel statistics given by 
%  the focal neighbourhood $F$ as follows:
%  
%  \begin{itemize}
%    
%    \item $N$ is a raster computed as a focal operation \FOCAL{R}{F}{sum}. Each
%    pixel represents the number of pixels from $R$ convoluted with the matrix
%    $F$. 
%    
%    \item $M$ is a raster computed as a focal mean \FOCAL{R}{F}{mean}, thus 
%each
%    pixel represents a mean value of its $F$-neighbourhood.
%    
%    \item $S$ is a raster computed as a focal standard deviation
%    \FOCAL{R}{F}{sd}, thus each pixel represents a standard deviation of its
%    $F$-neighbourhood.
%    
%  \end{itemize}
%  
%\end{definition}

@@

%\begin{definition}[Evaluation Run]
%  \label{def:EvalRun}
%  We define a single evaluation run as a tuple:
%  \begin{displaymath}
%    E = (V, m, p, w)
%  \end{displaymath}
%  \noindent where:
%  \begin{itemize}
%  
%    \item $V$ is the input dataset of points, representing the taxi dropoffs 
%in 
%    our case.
%    
%    \item $m$ is the metric used, in our case either \SOHUP or \SOHDOWN.
%    
%    \item $p$ represents the pixel size for aggregating points from $V$, 
%    e.g. $100 \times 100$ meters.
%    
%    \item $w$ represents the size of a weight matrix.
%    In our case, we chose a weight matrix depicted in 
%    Figure~\ref{fig:ExampleMatrices}(a) for both the G* and FocalG* cases.
%  \end{itemize}
%\end{definition}

%Figure~\ref{fig:ComparisonMorning} and Figure~\ref{fig:ComparisonEvening} show
%Standard and Focal $G^*$ computations for both morning and eventing datasets
%with weight matrix $W$ of size 3, 5, 7, 9, 15 and 31. In this figures, when 
%computing Focal
%$G^*$, the focal matrix $F$ has a constant size of $61{\times}61$ cells. 
%Example weight matrix $W$ and focal matrix $F$ are depicted in 
%Figure~\ref{fig:ExampleMatrices}.

Aggregation size 1 means we aggregate every point, which was in range of $100x100 \cdot 0.000001$ into one pixel. 
For aggregation size Z we aggregated ZxZ pixel from aggregated size 1 into a new pixel.
%
For our measurements we specified the following data series:
\begin{itemize}
 \item $W_i$: The weight matrix W is from 3 to 47 with a stepsize of 4.
 \item $F_j$: The focal matrix F is from  17 to 137 with a stepsize of 12. F is ignored for $G^*$ but not for Focal $G^*$
 \item $R_z$: The aggregation level is from 1 to 6 with a stepsize of 1. R represents the raster
\end{itemize}
\begin{definition}

  \begin{displaymath}
    \SOHUP(child, parent)
  \end{displaymath}
  \begin{displaymath}
    \SOHDOWN(child, parent)
  \end{displaymath}

  The SoH for a singled evalution run in the weight dimension is defined as:
  \begin{eqnarray*}
    child & = & G^*(R_z, W_{3+i \cdot 4}, st, F_{17+j \cdot 12}) \\
    parent & = & G^*(R_z, W_{3+(i+1) \cdot 4}, st, F_{17+j \cdot 12}) )
  \end{eqnarray*}
  The SoH for a singled evalution run in the focal dimension is defined as:
  \begin{eqnarray*}
    child & = & G^*(R_z, W_{3+i \cdot 4}, st, F_{17+j \cdot 12}) \\
    parent & = & G^*(R_z, W_{3+i \cdot 4}, st, F_{17+(j+1) \cdot 12}) )
  \end{eqnarray*}
  The SoH for a singled evalution run in the aggregation dimension is defined 
  as:
  \begin{eqnarray*}
    child & = & G^*(R_z, W_{3+i \cdot 4}, st, F_{17+j \cdot 12}) \\
    parent & = & G^*(R_{z+1}, W_{3+i \cdot 4}, st, F_{17+j \cdot 12}) )
  \end{eqnarray*}

  \begin{displaymath}
  z \in [1,6], i,j \in [0,10], G^* \in [Standard, Focal]
  \end{displaymath}
\end{definition}

Therefore we would calculate $10 \cdot 10 \cdot 5 + 10 \cdot 5=550$ results for each dimension.
The following conditions must hold $dim(F) > dim(W)$ and $dim(F) < dim(R)$ 
which leads e.g. to 460 results in the z-dimension.
We vary over the weight matrix W, the focal matrix F and the aggregation level, as motivated in the introduction.
To compare the impact each of these parameters has we compute all variations 
over two parameter and holding one parameter fix. We then calculate the mean as 
well as the standard deviation for the SoH for the fixed value based on the two 
other parameters. This allows us to isolate the impact of the variation in the 
single parameter. The results are then plotted in 
Figures~\ref{fig:SoHFocal}--\ref{fig:SoHZoom}, one image for each fixed 
parameter and the direction of the SoH.
The hotspots to use the SoH on are computed by the focal $G^*$ and standard $G^*$ method. The focal $G^*$ allows us, to isolate the impact that the size of the study area has on the stability of the results without recomputing the total dataset. 
\subsection{Clumping}

The SoH needs a parent and a child. Clusters can be calculated different. For
$G^*$ a sigmoid can be used. Values above 2.58 and under -2.58 means in less
then 1\% of the cases you are wrong. The values of Focal $G^*$ depend on the
focal size F. Therefore, we decide all values in the top quantile belong to a
clusters. For clustering we grouped the neighbourhood with values in the top
quantile to one cluster. Neigbours are are pixels with touching edges which
leads for every pixel to 8 neigbours. Fig.~\ref{fig:Clumping} depicts the 
clusters using different colours. In this figure, Focal~$G^*$ has smaller 
clusters which are distributed over the complete raster.

\begin{figure}[htp]
  \centering

  \begin{subfigure}{.7\linewidth}
    \caption{G* clustering difference}
    \includegraphics[width=\linewidth]{images/gen-metric-example-1}
    \label{fig:ClustDiffGstar}
  \end{subfigure}
  \hspace{3em}
  \begin{subfigure}{.7\linewidth}
    \caption{Focal G* clustering difference}
    \includegraphics[width=\linewidth]{images/gen-metric-example-2}
    \label{fig:ClustDiffFocalGstar}
  \end{subfigure}

  \caption{
    Example clustering for G* and Focal G* used as a input for SoH computation.
  }
  \label{fig:Clumping}
\end{figure}

\subsection{Aggregate}

The aggregation level is a slice plane through our three dimensional space. 
This is an example how the results can look with visualize representations of the calculated 
rasters and the raw data.
\begin{definition} Aggregation run:
\begin{eqnarray*}
    child & = & G^*(R_z, W_{11}, st, F_{41}) \\
    parent & = & G^*(R_{z+1}, W_{11}, st, F_{41})
  \end{eqnarray*}
\begin{displaymath}
z \in [1,6], G^* \in [Standard, Focal]
\end{displaymath}
\end{definition}

The results can be found in Figure~\ref{fig:Zoom}. It can be seen that with an
increase of the aggregation level the values increase. 
If the aggregation level increases the number of cluster decreases. It can seen that
small clusters are disappearing and some clusters are growing to big ones. From aggreagtion level~2 to
aggreagtion level~3 there is no big change in the images of Focal $G^*$ which is consitent with our results.
For the upward property, $G^*$ is always better than Focal
$G^*$, compared to the results from figure \ref{fig:SoHZoom}. Focal $G^*$
reaches the same result as $G^*$ at aggregation level~6 . For the downward
property, Focal $G^*$ leads to better results at aggregation level~6.

\begin{figure*}[htp]
  \centering
  \begin{tabular}{cccccccc}
    \multicolumn{8}{l}{
      Aggregation levels computed from original raster (log scale)
    }\\
    \hline
    1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 \\
    \includegraphics[width=4.6em]{images/gen-rawdata-1}&
    \includegraphics[width=4.6em]{images/gen-rawdata-2}&
    \includegraphics[width=4.6em]{images/gen-rawdata-3}&
    \includegraphics[width=4.6em]{images/gen-rawdata-4}&
    \includegraphics[width=4.6em]{images/gen-rawdata-5}&
    \includegraphics[width=4.6em]{images/gen-rawdata-6}&
    \includegraphics[width=4.6em]{images/gen-rawdata-7}&
    \includegraphics[width=4.6em]{images/gen-rawdata-8}
  \end{tabular}

  \vspace{1em}
  \begin{tabular}{cccccccc}
    \multicolumn{8}{l}{G* computed from multiple aggregation levels} \\
    \hline
    \includegraphics[width=4.6em]{images/gen-raw-zoom-gstar-1}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-gstar-2}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-gstar-3}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-gstar-4}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-gstar-5}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-gstar-6}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-gstar-7}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-gstar-8}\\

    \includegraphics[width=4.6em]{images/gen-demo-zoom-gstar-1}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-gstar-2}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-gstar-3}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-gstar-4}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-gstar-5}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-gstar-6}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-gstar-7}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-gstar-8}\\

  \end{tabular}

  \vspace{1em}
  \begin{tabular}{cccccccc}
    \multicolumn{8}{l}{Focal G* computed from multiple aggregation levels} \\
    \hline
    \includegraphics[width=4.6em]{images/gen-raw-zoom-focalgstar-1}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-focalgstar-2}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-focalgstar-3}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-focalgstar-4}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-focalgstar-5}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-focalgstar-6}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-focalgstar-7}&
    \includegraphics[width=4.6em]{images/gen-raw-zoom-focalgstar-8}\\

    \includegraphics[width=4.6em]{images/gen-demo-zoom-focalgstar-1}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-focalgstar-2}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-focalgstar-3}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-focalgstar-4}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-focalgstar-5}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-focalgstar-6}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-focalgstar-7}&
    \includegraphics[width=4.6em]{images/gen-demo-zoom-focalgstar-8}\\
  \end{tabular}

  \vspace{2em}
  \includegraphics[width=.45\linewidth]{images/gen-zoom-sohup-1}
  \hspace{1em}
  \includegraphics[width=.45\linewidth]{images/gen-zoom-sohdown-1}

  \caption{
    This image shows all aggregation levels for G* and Focal G* together 
    with two metrics -- $SoH^\uparrow$ and $SoH^\downarrow$.
  }
  \label{fig:Zoom}
\end{figure*}

\subsection{Blur}
We also made an example for the variation of the weight size W.
Blur is a slice plane through our three dimensional space. 
W is changed from 7 to 29 with a stepsize of 2, F and the aggregation level is fixed.
\begin{definition} Blur run:
\begin{eqnarray*}
    child & = & G^*(R_3, W_{7+i \cdot 2}, st, F_{41}) \\
    parent & = & G^*(R_3, W_{7+(i+1)\cdot 2}, st, F_{41})
  \end{eqnarray*}
\begin{displaymath}
i \in [0,11], G^* \in [Standard, Focal]
\end{displaymath}
\end{definition}
The blur results are plotted in Figure~\ref{fig:Blur}. The weight matrix is
shown in the first row, with increasing size. The 2nd and 4th row show how
$G^*$ and Focal $G^*$ changes based on W. This example is a slice plane of the
weight dimension ploted in figure ~\ref{fig:SoHWeight}. It can be seen that an
increase of the weight size leads in general to better results for $G^*$.
Also it can bee seen that for focal $G^*$ the clusters are more distributed over the raster.

\begin{figure*}[htp]
  \centering
  \begin{tabular}{cccccccc}
    \multicolumn{8}{l}{Multiple weight matrix sizes} \\
    \hline
    \includegraphics[width=4.6em]{images/gen-blur-wmatrices-1}&
    \includegraphics[width=4.6em]{images/gen-blur-wmatrices-2}&
    \includegraphics[width=4.6em]{images/gen-blur-wmatrices-3}&
    \includegraphics[width=4.6em]{images/gen-blur-wmatrices-4}&
    \includegraphics[width=4.6em]{images/gen-blur-wmatrices-5}&
    \includegraphics[width=4.6em]{images/gen-blur-wmatrices-6}&
    \includegraphics[width=4.6em]{images/gen-blur-wmatrices-7}&
    \includegraphics[width=4.6em]{images/gen-blur-wmatrices-8}
  \end{tabular}

@@

  \vspace{1em}
  \begin{tabular}{cccccccc}
    \multicolumn{8}{l}{Focal G* computed using different weight matrix sizes} \\
    \hline
    \includegraphics[width=4.6em]{images/gen-raw-blur-focalgstar-1}&
    \includegraphics[width=4.6em]{images/gen-raw-blur-focalgstar-2}&
    \includegraphics[width=4.6em]{images/gen-raw-blur-focalgstar-3}&
    \includegraphics[width=4.6em]{images/gen-raw-blur-focalgstar-4}&
    \includegraphics[width=4.6em]{images/gen-raw-blur-focalgstar-5}&
    \includegraphics[width=4.6em]{images/gen-raw-blur-focalgstar-6}&
    \includegraphics[width=4.6em]{images/gen-raw-blur-focalgstar-7}&
    \includegraphics[width=4.6em]{images/gen-raw-blur-focalgstar-8}\\

    \includegraphics[width=4.6em]{images/gen-demo-blur-focalgstar-1}&
    \includegraphics[width=4.6em]{images/gen-demo-blur-focalgstar-2}&
    \includegraphics[width=4.6em]{images/gen-demo-blur-focalgstar-3}&
    \includegraphics[width=4.6em]{images/gen-demo-blur-focalgstar-4}&
    \includegraphics[width=4.6em]{images/gen-demo-blur-focalgstar-5}&
    \includegraphics[width=4.6em]{images/gen-demo-blur-focalgstar-6}&
    \includegraphics[width=4.6em]{images/gen-demo-blur-focalgstar-7}&
    \includegraphics[width=4.6em]{images/gen-demo-blur-focalgstar-8}\\
  \end{tabular}

  \vspace{2em}
  \includegraphics[width=.45\linewidth]{images/gen-blur-sohup-1}
  \hspace{1em}
  \includegraphics[width=.45\linewidth]{images/gen-blur-sohdown-1}

@@

\section{Results and Discussion}

% Focal size  
%	Increase of focal size leads to better results in the upward property
% 	Standard deviation decrease with higher focal size 
%	No standard deviation af F 17x17 because of missing clusters
%	The downward property reaches its maximum at 65x65
%
%	Conclusion: F should be around 65x65 or higher
\subsection{Impact of study area}
The first parameter we want to isolate is the size of the study area. This enables us to discuss the optimal size of the study area for the given research/decision question. 
To do so, we have to fix the focal size F and use it as the x-axis and plot the 
mean value as well as the mean +/- the standard deviations of the SoH for all 
other parameter combinations as the y-axis.
In this scenario, only the Focal $G^*$ can produce varied results as the standard $G^*$ has by definition a fixed study area by always using the total size of the study area. 
In Figure~\ref{fig:SoHFocal} the \SOHUP and \SOHDOWN is shown for $G^*$ and 
Focal $G^*$.
Higher values indicate that the hot spots found are more stable. 
The \SOHUP in Figure~\ref{fig:fUp} growth with an increase of the focal size.
It can see that an increase of the focal size leads to better results for \SOHUP. 
With a higher focal size you can see that the dotted line, which represent the standard deviation, is getting smaller. This indicates that an increase in the size of the study area does not only lead to an overall increase of stability but also that the it reduces the volatility of the variations in other parameters.
When F and W have similar size, there are no clusters. Because the values have no much variation. 
\SOHDOWN (Figure ~\ref{fig:fDown}) shows a different picture then the \SOHUP.
It reaches its maximum SoH for a focal size F of 65x65 in our evaluation set. 
The standard deviation is not decreasing with an increase of the 
focal size.

\begin{figure}[htp]
  \begin{subfigure}{\linewidth}
    \caption{SoH up for focal size}
    \includegraphics[width=\linewidth]{images/gen-focal-1}
    \label{fig:fUp}
  \end{subfigure}
  \hspace{1em}
  \begin{subfigure}{\linewidth}
    \caption{SoH down for focal size}
    \includegraphics[width=\linewidth]{images/gen-focal-2}
    \label{fig:fDown}
  \end{subfigure}
  \caption{SoH for focal matrix size}
  \label{fig:SoHFocal}
\end{figure}

>>>>

\begin{abstract}
Hot spot analysis is essential for geo-statistics.

\end{abstract}

@@

Huge in increase in data, e.g. even without incorporating spatial structures the OSM\cite{OpenStreetMap} data, as example for VGI, is over 1 TB

\subsection{Examples}

Spatio-temporal sentiment hotspot detection using geotagged photos \cite{Zhu:2016:SSH:2996913.2996978}

A framework for evacuation hotspot detection after large scale disasters using location data from smartphones: case study of Kumamoto earthquake \cite {Yabe:2016:FEH:2996913.2997014}

\cite{Wiener:2016:BCR:2996913.2996931} BigGIS Vision Paper

\cite{7129542} The era of big spatial data : 
"Unfortunately, the urgent need to manage and analyze big
spatial data is hampered by the lack of specialized systems,
techniques, and algorithms to support such data." "We
provide general guidelines to show the four components that
are needed in a system for big spatial data." 
In general an overview of approaches in 2015 and overall targets for systems

\cite{mehta2016spatio}   Spatio-Temporal Hotspot Computation on Apache Spark (GIS Cup)

Giscup 2016 \footnote{http://sigspatial2016.sigspatial.org/giscup2016/home}

\section{State of the Art}
\subsection{Geostatistic}
While the standard hot spot analysis approaches allow for a fast and automated exploratory analysis of spatial data sets, they are highly dependent on their parametrizations and underlying data. This is quite similar to the challenges in clustering, in particular for the well-known k-means, first coined in \cite{macqueen1967some}, or DBScan (\cite{ester1996density}) algorithm. This leads to an instability of the analysis results, which are then difficult to use with high certainty. Here, a selection of approaches and discussions of the last 15 years are presented how to tackle this instability.

The most general approach is to pre-determine and calculate spatial dependencies, based on the a-priori knowledge of the analyst, the empirical data set or best-practices from previous, similar analyses. A good example for this approach can be found in \cite{suomi2012effects}. In their work, the authors examine the effects of scale on temperature and in particular urban heat island modeling. The pre-determined range of spatial influences is called \emph{buffer zones}; these buffer zones indicate the range of impact in an inner-city temperature measurement scenario. They found that for their empirical data set a buffer zone of 1000m provides the best results. This is familiar to the well-known approach to solve the inherent problem of DBScan, the determination of its distance: OPTICS (\cite{ankerst1999optics}). The disadvantage of this approach is that it has to be manually pre-determined, which results in similar problems as the semi-supervised methods to detect points of interest. It can not be done automatically without introducing an element of uncertainty, which is opposed to our goal to create a more stable approach.

An automated approach is presented with the \emph{A Multidirectional Optimal Ecotope-Based Algorithm} (AMOEBA) in \cite{aldstadt2006using}. The idea behind this approach is to automatically create the optimal, scale-invariant weight matrix and then use this weight matrix in conjunction with a clustering approach to create a graphical overview map of areas of interest. The term ecotope is used for this areas, which is the technical term from the field of biology for the habitat of species. The result is a consistent identification of spatial clusters on a map. In their work they use the $G^*$ statistic as the underlying statistic. The clustering approach is quite similar to DBScan in its approach of creating ecotopes.

A true modification of the $G^*$ statistic is presented in a later work (\cite{getis2010constructing}) of the same authors called the LSM (local statistics model). They base their modification on the Kriging approach and its ability to model the spatial autocorrelation as a function dependent on the distance. The idea is to model the weight matrix W as a function of the spatial autocorrelation, where each entry of the matrix is a value derived from the empirical  (semi-)variogram. This leads to continuous values up to the so called \emph{critical distance}, which is "defined as the distance beyond which no discernible increase in clustering exist" (\cite{getis2010constructing}). They compare their configuration to other, well-known spatial configuration approaches for the weight matrix W. These are taken from \cite{griffith1996some} and in the words of \cite{getis2010constructing}:
''Research on W has been reviewed by Griffith (1996, p. 80), who concludes that five rules of thumb aid in the specification of weights matrices: 
\begin{enumerate}
\item “It is better to posit some reasonable geographic weights matrix than to assume independence.” This implies that one should search for or theorize about an appropriate W and that better results are obtained when distance is taken into account.
\item “It is best to use surface partitioning that falls somewhere between a regular square and a regular hexagonal tessellation.” Griffith suggests that for planar data, a specification between four and six neighbors is better than something either above six or below four. Of course, the configuration of the planar tessellations will play a role here (\cite{boots2000global}).
\item “A relatively large number of spatial units should be employed, n > 60.” Following from the law of large numbers, most spatial research, especially due to unequal size spatial units, would require fairly large samples.
\item “Low-order spatial models should be given preference over higher-order ones.” Following from the scientific principle of parsimony, it is always wise to choose less complicated models when the opportunity presents itself.
\item “In general, it is better to apply a somewhat under-specified (fewer neighbors) rather than an over-specified (extra neighbors) weights matrix.” \cite{florax1995impacts} found this result by identifying the power of tests. Overspecification reduces power. They recognize that “Uncertainty with respect to proper specification has long been recognized as a fundamental problem in applied spatial econometric modeling” (p. 132).
\end{enumerate}
''

\cite{ord2001testing} discuss the question in how to formulate the $G^*$ statistic to focus more on local pattern, while still accounting for the global autocorrelation. They propose the $O$ statistic which uses the (semi-)variogram to subdivide the data set into several '' 'relatively homogeneous' subregions'' (\cite{ord2001testing}). This allows the identification of smaller, more local hot spots, which can be overshadowed in bigger data sets. Finally, they restrict the general applicability in that the version presented in their work requires spatial stationarity.

\cite{westerholt2015local} present a scale-sensitive version of the local $G^*$ statistic, which they call the \emph{GS statistic}. The motivation for this modification is to account for the differences in the scale (the impact of the area under investigation) of the data set, i.e. whether a data set includes only the inner city or also its surrounding area. The problem lies in the detection and use of the local context of the gathered data. A fixed weight matrix W does not include the difference in context, e.g. in Twitter feeds. Their approach is to redefine the neighborhood of a data point with upper and lower distance thresholds, which are then used in pairwise comparisons. Only sufficiently connected data points within their thresholds are considered to be viable as a hot spot and only those points are used for the global mean and global deviation values. They evaluated their approach on Twitter data of the city of San Francisco, USA and show that this leads to reduction or even negation of cross-scale interference. The main restrictions of this approach lies in its increased computational costs as well as its reliance on continuous distance functions.

For futher literature regarding the creation of optimal weight matrices for spatial associations we refer to the work of \cite{aldstadt2006using}, where they provide an exhaustive overview of the state of the art.

\cite{Bruns2017} for focal $G^*$ and efficient parametrization of $G^*$

\subsection{Frameworks and Methods}
An Effective High-Performance Multiway Spatial Join Algorithm with Spark \cite{ijgi6040096}
"In this paper, we present an effective
high-performance multiway spatial join algorithm with Spark (MSJS) to overcome the multiway
spatial join bottleneck." "Spatial analysis is the core engine that drives research and applications of GIS."

GeoSpark: a cluster computing framework for processing large-scale spatial data \cite{Yu:2015:GCC:2820783.2820860}

Spatio-Temporal Join on Apache Spark \cite{Whitman:2017:SJA:3139958.3139963}

Algorithmically-Guided User Interaction \cite{vanDijk:2017:AUI:3139958.3140032}

GeoMesa: a distributed architecture for spatio-temporal fusion \cite{hughes2015geomesa}

geotrellis\footnote{https://geotrellis.io/}

@@

\subsection{Hot Spot Analysis}
The goal of hotspot analysis is the detection of interesting areas as well as patterns in spatial information. One of the most fundamental approach is Moran´s I~\cite{MoranI}. There it is tested whether or
not a spatial dependency exists. This gives the information on global
dependencies in a data set. Upon this hypothesis test several geo-statistical
tests are based. The most well known are the Getis-Ord statistic~\cite{Ord.1995}
and LISA~\cite{Anselin.1995}. In both cases the general, the global statistic of
Moran´s I is applied in a local context. The goal is to detect not only global
values, but instead to focus on local hotspots and to measure the significance
of those local areas.

Existing methods for determining hot spots are dependent on the parametrization
of the weight matrix as well as on the size of the study area. Intuitively,
increasing the size of a weight matrix has a "blurring" effect on the raster
(

@@

\begin{figure*}[htp]
  \centering

@@

@@
